Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-09-13 article.tex
Update

\paragraph{Case 4: Inflation of an idealised ventricle}
Inflation of an idealised ventricle (Figure \ref{fig:ventricle}) was proposed by \citet{Land2015} as a benchmark problem for cardiac mechanics software.
The case is 3-D, static, with finite hyperelastic strains.
The initial geometry is defined as a truncated ellipsoid:
\begin{eqnarray}
	x = r_s \sin(u) \cos(v), \quad
	y = r_s \sin(u) \sin(v), \quad
	z = r_l \cos(u)
\end{eqnarray}
where on the inner (endocardial) surface $r_s =7$ mm, $r_l = 17$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{17} \right) \right]$ and $v \in \left[-\pi, \pi \right]$, while on the outer (epicardial) surface $r_s =10$ mm, $r_l = 20$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{20} \right) \right]$ and $v \in \left[-\pi, \pi \right]$.
The base plane $z = 5$ mm is implicitly defined by the ranges for $u$.
%Constitutive parameters: isotropic, 
The hyperelastic material behaviour is decsribed by the transversely isotropic constitutive proposed by \citet{Guccione} law, where the parameters $C = 10$ kPa, $b_f = b_t = b_{fs} = 1$.
The chosen parameters produce isotropic behaviour.
A pressure of 10 kPa is applied to the inner surface, and the base plane is fixed.
The geometry is meshed using structured approach and is predominantly composed of hexahedra, with prism cells forming the apex.
Four succesively refined meshes are examined: 1,620 (shown in Figure  \ref{fig:ventricle}), 12,960, 103,680, and 829,440 cells.
%Figure \ref{fig:ventricle} shows the mesh with 1,620 cells (1,512 hexahedra and 108 prisms).
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.4\textwidth]{figures/ventricle} 
   \caption{Idealised ventricle case geometry, showing the mesh with 1,620 cells}
   \label{fig:ventricle}
\end{figure}

\paragraph{Case 5: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
In current work, the finite strain elastoplastic version \citep{Simo1992} is considered. \hl{We could also consider the small-strain elastic version if needed}
The 2-D tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The prescribed shear traction is $\tau = 0.3125$ MPa.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
There are no body forces, and the problem is solved as statically as 2-D plane strain.
The problem is solved as static, using 30 equally-sized loading increments.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry.png} 
   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
   \label{fig:cooks_membrane}
\end{figure}

@@

The predicted $\sigma_{xx}$ stress distribution for hexahedral mesh with $512\,000$ cells is shown in Figure \ref{fig:mms_stress}(a).
The corresponding cell-wise $\sigma_{xx}$ error distribution is shown in Figure \ref{fig:mms_stress}(b), where the errors of greatest magnitude ($-29$ kPa) occur at the ends of the boundary.
\begin{figure}[htbp]
	\centering
	\subfigure[Predicted $\sigma_{xx}$ stress distribution]
	{
		\label{fig:mms_sxx}
   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx} 
   	}
	\subfigure[Cell-wise $\sigma_{xx}$ error distribution]
	{
		\label{fig:mms_sxx_diff}
   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx_diff}  
   	}
	\caption{Manufactured solution cube case: the predicted $\sigma_{xx}$ stress distribution on a cut-plane for the hexahedral mesh with $512\,000$ cells (left).}
	\label{fig:mms_stress}
\end{figure}

Figure \ref{fig:mms_accuracy}(a) shows the displacement magnitude discretisation errors ($L_2$ and $L_\infty$) as a function of the average cell width for both the hexahedral and polyhedral meshes.
Figure \ref{fig:mms_accuracy}(b) shows the corresponding order of accuracy plots.
The maximum ($L_\infty$) and average ($L_2$) discretisation errors are seen to reduce at an approximately second-order rate for both hexahedral and polyhedral meshes, where the errors are smaller on the hexahedral meshes \hl{check}.
The discretisation errors in the $\sigma_{xx}$ component of stress are shown in Figure \ref{fig:mms_accuracy}(b), and the order of accuracy in \ref{fig:mms_accuracy}(d).
The order of accuracy for the average stress magnitude error is seen to be \hl{XX} for the hexahedral meshes and just over 1.5 for the polyhedral meshes.
In contrast, the maximum stress error order of accuracy is seen to be 1 for both mesh types \hl{check}.
\begin{figure}[htbp]
	\centering
	\subfigure[Displacement magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_dispErrors} 
   	}
	\subfigure[Displacement discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_disp_orderOfAccuracy}  
   	}
	\subfigure[Stress magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_stressErrors} 
   	}
	\subfigure[Stress discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_stress_orderOfAccuracy}  
   	}
	\caption{Manufactured solution cube case: the accuracy and order of accuracy for displacement and stress}
	\label{fig:mms_accuracy}
\end{figure}

The presented results have been generated using the Jacobian-free Newton-Raphson solution algorithm; however, minimal differences were seen when using the segregated solution algorithm, and hence, the results are not shown \hl{check}.
Analysis of time and memory requirements is left to subsequent sections.

>>>>

\paragraph{Case 4: Inflation of an idealised ventricle}
Inflation of an idealised ventricle (Figure \ref{fig:ventricle}) was proposed by \citet{Land2015} as a benchmark problem for cardiac mechanics software.
The case is 3-D, static, with finite hyperelastic strains.
The initial geometry is defined as a truncated ellipsoid:
\begin{eqnarray}
	x = r_s \sin(u) \cos(v), \quad
	y = r_s \sin(u) \sin(v), \quad
	z = r_l \cos(u)
\end{eqnarray}
where on the inner (endocardial) surface $r_s =7$ mm, $r_l = 17$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{17} \right) \right]$ and $v \in \left[-\pi, \pi \right]$, while on the outer (epicardial) surface $r_s =10$ mm, $r_l = 20$ mm, $u \in \left[-\pi, -\arccos \left( \frac{5}{20} \right) \right]$ and $v \in \left[-\pi, \pi \right]$.
The base plane $z = 5$ mm is implicitly defined by the ranges for $u$.
%Constitutive parameters: isotropic, 
The hyperelastic material behaviour is described by the transversely isotropic constitutive proposed by \citet{Guccione} law, where the parameters $C = 10$ kPa, $b_f = b_t = b_{fs} = 1$.
The benchmark specifies an incompressible material; in the current work, incompressibility is enforced weakly using a penalty approach, where the bulk modulus parameter is chosen to be two orders of magnitude greater ($\kappa = 1000$ kPa) than the greatest shear modulus parameter.
The chosen parameters produce isotropic behaviour.
A pressure of 10 kPa is applied to the inner surface, and the base plane is fixed.
The geometry is meshed using a structured approach and is predominantly composed of hexahedra, with prism cells forming the apex.
Four successively refined meshes are examined: 1,620 (shown in Figure  \ref{fig:ventricle}), 12,960, 103,680, and 829,440 cells.
%Figure \ref{fig:ventricle} shows the mesh with 1,620 cells (1,512 hexahedra and 108 prisms).
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.4\textwidth]{figures/ventricle} 
   \caption{Idealised ventricle case geometry, showing the mesh with 1,620 cells}
   \label{fig:ventricle}
\end{figure}

\paragraph{Case 5: Cook's membrane}
Cook's membrane (Figure \ref{fig:cooks_membrane}) is a well-known bending-dominated benchmark case used in linear and non-linear analysis.
The current work considers the finite strain elastoplastic version \citep{Simo1992}. \hl{We could also consider the small-strain elastic version or hyperelastoplastic version}
The 2-D tapered panel (trapezoid) is fixed on one side and subjected to uniform shear traction on the opposite side.
The prescribed shear traction is $\tau = 0.3125$ MPa.
The vertices of the trapezoid (in mm) are (0, 0), (48, 44), (48, 60),  and (0, 44).
There are no body forces, and the problem is solved as statically as 2-D plane strain.
The problem is solved as static, using 30 equally-sized loading increments.
\begin{figure}[htbp]
   \centering
   \includegraphics[width=0.4\textwidth]{figures/cooksMembrane-geometry.png} 
   \caption{Cook's membrane case geometry (dimensions in mm) and loading conditions (fixed left, traction $\tau$ applied to right)}
   \label{fig:cooks_membrane}
\end{figure}

@@

Figure \ref{fig:mms_disp_accuracy}(a) shows the displacement magnitude discretisation errors ($L_2$ and $L_\infty$) as a function of the average cell width for both the hexahedral and polyhedral meshes.
Figure \ref{fig:mms_disp_accuracy}(b) shows the corresponding order of accuracy plots.
For ease of interpretation, the symbol shapes in the figures have been chosen to correspond to the cell shapes: a square for hexahedra, a triangle for tetrahedra and a pentagon for polyhedra.
The maximum ($L_\infty$) and average ($L_2$) discretisation errors are seen to reduce at an approximately second-order rate for all mesh types, except for the $L_2$ error on the hexahedral meshes, which is approximately 2.3 on the finest grid.
%where, interestingly, the errors are, on average, smaller on the polyhedral meshes.
%The reason for this is unclear but may be due to the stencil of the structured polyhedra being more isotropic and containing close neighbours.
\begin{figure}[htbp]
	\centering
	\subfigure[Displacement magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_dispErrors} 
   	}
	\subfigure[Displacement discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_disp_orderOfAccuracy}  
   	}
	\caption{Manufactured solution cube case: the accuracy and order of accuracy for displacement magnitude}
	\label{fig:mms_disp_accuracy}
\end{figure}

The predicted $\sigma_{xx}$ stress distribution for a hexahedral mesh with $512\,000$ cells is shown in Figure \ref{fig:mms_stress}(a).
The corresponding cell-wise $\sigma_{xx}$ error distribution is shown in Figure \ref{fig:mms_stress}(b).
The errors of greatest magnitude ($43$ kPa) occur at the boundaries, corresponding to where the local truncation error is higher due to the use of one-sided differencing.
\begin{figure}[htbp]
	\centering
	\subfigure[Predicted $\sigma_{xx}$ stress distribution]
	{
		\label{fig:mms_sxx}
   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx} 
   	}
	\subfigure[Cell-wise $\sigma_{xx}$ error distribution]
	{
		\label{fig:mms_sxx_diff}
   		\includegraphics[height=0.35\textwidth]{figures/mms_sxx_diff}  
   	}
	\caption{Manufactured solution cube case: the predicted $\sigma_{xx}$ stress distribution on a cut-plane for the hexahedral mesh with $512\,000$ cells (left).}
	\label{fig:mms_stress}
\end{figure}

The discretisation errors in the stress magnitude are shown in Figure \ref{fig:mms_stress_accuracy}(a), and the order of accuracy in Figure \ref{fig:mms_stress_accuracy}(b).
The order of accuracy for the average ($L_2$) stress error is seen to be approximately 1.9 for the hexahedral meshes and 1.5 for the polyhedral meshes.
Unlike the displacement errors where the order of accuracy for the average and maximum errors are similar, the maximum ($L_\infty$) stress order of accuracy is noticeably lower and is seen to approach 1 for both mesh types.
\begin{figure}[htbp]
	\centering
	\subfigure[Stress magnitude discretisation errors]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_stressErrors} 
   	}
	\subfigure[Stress discretisation error order of accuracy]
	{
   		\includegraphics[width=0.45\textwidth]{figures/mms_stress_orderOfAccuracy}  
   	}
	\caption{Manufactured solution cube case: the accuracy and order of accuracy for stress magnitude}
	\label{fig:mms_stress_accuracy}
\end{figure}

The presented results have been generated using the Jacobian-free Newton-Raphson solution algorithm; however, minimal differences were seen when using the segregated solution algorithm, and hence, the results are not shown.
% Note: this is PETSc segregated so discretisation is exactly the same; OF segregated will be slightly different as boundary gradients are different
% \hl{check - maybe one comparison graph}.
%\hl{Seg and JFNK are the same when including boundary values but this causes a problem for the poly grid}

Analysis of time and memory requirements is left to the subsequent section.

%%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Where do we put accuracy results?}
%%%--------------------------------------------------------------------------------------------------------------------%%
\hl{Where do we put accuracy comparison plots?}
\hl{Maybe we should change the previous section to be accuracy tests instead of just order of accuracy}
\hl{For now, put results here}

The prediction deformed configuration of the midline of the ventricle wall is shown in Figure \ref{fig:ventricle_accuracy}, where a side-by-side comparison is given with the round-robin benchmark results from \citet{Land2015}.
The predictions of the current work fall within the ranges of the benchmark.
\begin{figure}[htbp]
	\centering
	\subfigure[Predictions from the current work]
	{
   		\includegraphics[width=0.46\textwidth]{figures/ventricle_results} 
   	}
	\subfigure[\citet{Land2015} benchmarks]
	{
   		\includegraphics[width=0.48\textwidth]{figures/ventricle_results_land2015}  
   	}
	\caption{Idealised ventricle: predictions for the current method (left) vs the round-robin predictions from \citet{Land2015}}
	\label{fig:ventricle_accuracy}
\end{figure}
